---
title: "STAT 545A Homework 5"
author: "Tian Gao"
date: "2018/9/27"
output: github_document
---

# Bring rectangular data in
```{r}
## load gapminder
suppressPackageStartupMessages(library(gapminder))
## load tidyverse
suppressPackageStartupMessages(library(tidyverse))
## load forcats
suppressPackageStartupMessages(library(forcats))
## load plotly
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))
library(gapminder)
library(knitr)
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(plotly))
```

# Part 1: Factor management

* Elaboration for the gapminder data set
* Drop Oceania. Filter the Gapminder data to remove observations associated with the continent of Oceania. Additionally, remove unused factor levels. Provide concrete information on the data before and after removing these rows and Oceania; address the number of rows and the levels of the affected factors.

* First let's see the levels of gapminder.
```{r}
levels(gapminder$continent)
```

* Then I'll show the rows of a gapminder
```{r}
nrow(gapminder) 
```

* Then let's drop Ocania
```{r}
drop <- gapminder %>% 
  filter(continent %in%  c("Europe", "Africa", "Asia", "Americas"))
levels(drop$continent)
```

* We can see that the continent of Oceania is unchanged. To fix this we need to use droplevels()
```{r}
dropped <- drop %>%
  droplevels()

levels(dropped$continent)
```

* Then I'll address the number of rows and the levels of continent before and after removing Oceania.
```{r}
# before 
nrow(gapminder)
nlevels(gapminder$continent)
# after
nrow(dropped)
nlevels(dropped$continent)
```
* From the above discussion, we can see that the number of rows before dropping Oceania is 1704 and the continent level is 5, the number of rows after dropping oceania is 1680 and continent level is 4.

* Reorder the levels of country or continent. Use the forcats package to change the order of the factor levels, based on a principled summary of one of the quantitative variables. Consider experimenting with a summary statistic beyond the most basic choice of the median.
 * First of all let's show the dropped gapminder
```{r}
dropped %>%
  head(15) %>%
  knitr::kable()
```

* We can see that the data frame is ordered by country in alphabetical order. 

* I'll take a look on asia, in 1970
```{r}
asia1972 <- gapminder %>%
  filter(year == 1972, continent == "Asia")

asia1972 %>%
  head(20) %>%
  kable()
```

* Here I want to arrange the table ordered by gdpPerCap
```{r}
asia1972 %>%
  arrange(gdpPercap) %>%
  mutate(country = fct_reorder(country, gdpPercap)) %>%
  head(20) %>%
  kable()
```

```{r}
asia1972 %>%
  arrange(gdpPercap) %>%
  mutate(country = fct_reorder(country, gdpPercap)) %>%
  # show preview of resultant levels
  ggplot(aes(x = gdpPercap, y=country), color = country) +
  # make it a scatterplot
  geom_point() + 
  # scale x axis by log10
  scale_x_log10() +
  # change axis labels
  xlab("LifeExpectency") +
  ylab("country") +
  # change theme
  theme_bw()
```
* Conclusion, We need both arrange for table and fct_roder for plot

# Part 2: File I/O
* Experiment with one or more of write_csv()/read_csv() (and/or TSV friends), saveRDS()/readRDS(), dput()/dget(). Create something new, probably by filtering or grouped-summarization of Singer or Gapminder. I highly recommend you fiddle with the factor levels, i.e. make them non-alphabetical (see previous section). Explore whether this survives the round trip of writing to file then reading back in.

* In this part I'll use the data from last question. 
* Let's check the I/O
* write_csv()/read_csv()
```{r}
write_csv(asia1972,"gap_asia1972.csv")

read_csv("gap_asia1972.csv") %>%
  head(20) %>%
  knitr::kable()
```

* saveRDS()/readRDS()
```{r}
# save to rds
saveRDS(asia1972,"gap_asia1972.rds")

# read from rds
readRDS("gap_asia1972.rds") %>%
    head(20) %>%
  knitr::kable()
```

* dput()/dget()
```{r}
# save to rds
dput(asia1972,"gap_asia1972.txt")

# read from rds
dget("gap_asia1972.txt") %>%
    head(20) %>%
  knitr::kable()
```

* conclusion: Using CSV format, R is required to parse clumn to default format. The data however keeps unchanged. saveRDS and readRDS can keep both data and classes of each column.dput() and dget() can keep both data and classes of each column. 

# Part 3: Visualization design
* Remake at least one figure or create a new one, in light of something you learned in the recent class meetings about visualization design and color. Maybe juxtapose your first attempt and what you obtained after some time spent working on it. Reflect on the differences. If using Gapminder, you can use the country or continent color scheme that ships with Gapminder. Consult the dimensions listed in All the Graph Things.
* For this part I'll remake a figure. 

```{r}
gapminder %>%
  filter(continent == "Americas") %>%
  ggplot(aes(x = lifeExp, y = gdpPercap)) +
  geom_point() +
  scale_y_log10()
```

* From the  above plot we can see the rough relation between lifeExp and gdpPercap. 
* I'll try to edit a new version to reflect more dimensions other than lifeExp and gdpPercap

```{r}
gapminder %>%
  filter(continent == "Americas") %>%
  ggplot(aes(x = lifeExp, y = gdpPercap, colour = country)) +
  geom_point() +
  scale_y_log10(label = dollar_format())+
  theme_bw() + # change theme
  ylab("GDP per capita") + # change title on Y axis
  xlab("Life Expectancy") #change title on X axis
```
* This is the new version, I'll add another dimension to the plot which is the country. From the new version, we can see clearly that the trend of GDP per capita versus Life Expectancy in different countries. Also adding the dollar sign could make it more readable for us.

* Then, make a new graph by converting this visual (or another, if you¡¯d like) to a plotly graph. What are some things that plotly makes possible, that are not possible with a regular ggplot2 graph?

* For this one I'll convert the above plot to plotly. 
```{r}
plot <-gapminder %>%
  filter(continent == "Americas") %>%
  ggplot(aes(x = lifeExp, y = gdpPercap)) +
  geom_point() +
  scale_y_log10()

ggplotly(plot)
```
* From the plotly above, we can see that when I can hover over the points to see the detailed information of that point.
* Also, I can zoom in by dragging a box in the plot. 
```{r}
gapminder %>%
  filter(continent == "Americas") %>%
  plot_ly(x = ~lifeExp,
          y= ~gdpPercap, 
          z= ~country, 
          color = ~country,
          type = "scatter3d",
          mode = "markers") %>%
    layout(scene = list(xaxis = list(type = "log"))) #log z
```

* The above plot spread the 2-demension version to a 3-demension one. Once again, we can see that we can interact with the point here and drag the plot to view it in different angles. 


# Part 4: Writing figures to file
Use ggsave() to explicitly save a plot to file. Then use  to load and embed it in your report. You can play around with various options, such as:

* Arguments of ggsave(), such as width, height, resolution or text scaling.
Various graphics devices, e.g. a vector vs. raster format.
Explicit provision of the plot object p via ggsave(..., plot = p). Show a situation in which this actually matters.
